MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites

### 2014-09-15 chapters/MetaTracts.tex
minor changes
updated images
updated text

Carbon fiber reinforced polymer components, more specifically CFRP laminates with endless carbon fibers which we focus on in this work, typically consist of two main components: 
The matrix on the one side, which acts as a bonding component and the reinforcements on the other side, which allows for achieving the desired characteristics. To manufacture CFRP laminates with endless carbon fibers, various production processes are in industrial use. Most of these processes start with the reinforcement component by weaving individual carbon fiber bundles (yarn) into sheets of carbon fiber cloth in a predefined pattern. These sheets of woven carbon fiber cloth, which form the geometrical structure of the final CFRP materials, are also refereed to as fabric. Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth. Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the strength, durability and stiffness properties. Next, resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component. The main function of this matrix component is to act as a bonding component between the individual carbon fiber bundles. For the matrix of CFRP material systems with endless carbon fibers typically resins such as epoxy are frequently used. In special applications additional filler materials, such as colors or other particles, may be integrated. The production process CFRP laminate is finished after curing the resin and the CFRP laminate is ready for further processing.

@@

While fiber bundles are now understood as highly important in determining component properties, the tools for visualization of the internal structure have not developed at the same pace.
To the best of our knowledge, there is no current work that can resolve simple queries such as:
\begin{itemize}
\item{ How to extract and visualize the geometric structure of a particular fiber bundle?}
\item{ How to visualize the interaction between a particular pair of fiber bundles (crossovers/braiding) or a unit cell?}
\item{ Which fiber bundles show a particular orientation, how may different orientations are prevalent in a unit cell? }
\item{ Which fiber bundles are of the same type of yarn (i.e. which bundles show similar sizes or diameters), which is the largest or smallest fiber bundle?}
\end{itemize}

@@

\subsection {Fiber Clustering in Diffusion Tensor Imaging}
\label {subsec:fiberClus}
%!!!!! TODO REFINE FIBER CLUSTERING !!!!!
%The next step is to identify  anatomically relevant fiber bundles. This is achieved through fiber clustering.
The general assumption of fiber clustering in diffusion tensor imaging is that a similarity measure based on factors such as proximity between fibers can be used to interpret a collection of fiber tracts as bundles. Extensive research has been done on automatic DTI fiber clustering methods (see \cite{Corouge2004},~\cite{Brun2003},~\cite{Brun2004},~\cite{Zhang2008},~\cite{westinMEDIA02}).

@@

Figure \ref{fig:data-char} shows dataset1 with woven fiber bundles. Dataset1 is 450x300x500 voxels in size with isotropic resolution of $2\mu m$, the datatype is uint8. It clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern. The unit cell is used as a basis structure for CFRP manufacturing. The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material. Figure~\ref{fig:data-char}A shows a volume rendering of the dataset. B shows 2D slices along the X(right) and C in Z(left) axis. Figure~\ref{fig:data-char}D shows the zoomed in version of the green square. 
The resolution of dataset1 renders the individual carbon fibers in a fiber bundle are hard to resolve using 2D slice images only. Figure~\ref{fig:data-char}D contains multiple bundles going in opposite directions. However, the separation between two fiber bundles is barely visible. The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization. The fiber bundles themselves may differ in terms of the amount of fibers in the bundle, figure~\ref{fig:data-char}A shows the large variation in cross section sizes among the bundles.
Depending on the weaving pattern the fiber bundles cross each other in different orientations. In general the number of orientations is limited by the weaving pattern. In all our data sets the number of separate orientations was two. Nevertheless, the weaving pattern may cause the individual bundles to be curved. In consequence of the weaving pattern of the fiber bundles, individual fibers may be adjacent in euclidean distance but belong to different bundles. 
We make the following assumptions on our data:
\begin{itemize}
\item The structure embedded in the data contains fiber bundles of indiscernible fibers.
\item (Local orientation) Each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
\item The local orientation may change gradually inside the fiber bundle.
\item The local orientation may be noisy and not reliable.
\item (Connectivity) Moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
\item Fiber bundles going in different directions only interact near the surface of contact.
\end{itemize}
%\begin{figure}[htp]
%  \centering
%    \subfloat[Overview of the dataset]{\label{figur:3}\includegraphics[scale=0.3]{imagesMetaTracts/crop13-a.eps}}
%  \subfloat[X Axis]{\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/xaxis.eps}}\label{figur:1}\\
%  \subfloat[Y Axis]{\label{figur:2}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/yaxis.eps}}
%  \subfloat[Z Axis]{\label{figur:3}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/zaxis.eps}}
%  \caption{One of our datasets.Fig \ref{fig:orig_data}(a) provides an overview of the entire volume, the rest show slices of each of the major axis.}\label{fig:orig_data}
%\end{figure}

@@

\begin{figure}[tb]
%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
\includegraphics[width=0.5\textwidth]{imagesMT2014/image3.png}
\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
\label{fig:flowchart}
\end{figure}

Figure \ref{fig:flowchart} shows our pipeline. The approach consists of multiple steps which we divide into two broad phases, extracting the MetaTracts  (this section) and  finding plausible fiber bundles (Section~\ref{subsec:fiber-bundles}).
%\subsection {Extracting MetaTracts}
%\label{subsec:ExFiberCyl}
%We explain the different steps of MetaTracts extraction.

@@

\subsection {MetaTracts generation}
\label{subsec:MetaTracts-generation}
In $\mathbb{R}^2$ all the above properties hold except that cylinders are replaced by rectangles. We explain the process of MetaTracts generation in $\mathbb{R}^2$. The procedure trivially extends to $\mathbb{R}^3$. The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a $R_H$[0,1]. In Figure \ref{fig:algo} all regions which have reliable Hessian less than the threshold have been marked in green. Let the seed point associated with the MetaTract be grid point $C_{p}$ (Property 2, Section~\ref{subsec:mtprop}). The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the blue arrow (Figure~\ref{fig:algo}C). We generate a rectangle of length L and radius R, the blue dashed rectangle shows the rectangle generated at $C_p$ (Property 3 and 5, Section~\ref{subsec:mtprop}). 
\begin{figure}[tb]
  \begin{center}
   \includegraphics[ width=0.45\textwidth]{imagesMT2014/image4}
 % \includegraphics[ width=0.45\textwidth]{imagesMT2014/MT_abstract} 
  \end{center}
 \caption{MetaTracts generation: (B) MetaTract generation, region with $R_H$ less than threshold. (A) close up of marked region in (B). (C) Parts of a single MetaTract.}
  \label{fig:algo}
\end{figure}
The set of vertices which are in the blue region but not overlapped by green are possible candidates for the start point of the next cylinder in a particular MetaTract. We call these vertices \textit{``candidate vertices"}. From these we select another grid vertex which will be the start point for the next cylinder.
%\begin{figure}
%\centering
% \includegraphics[width=0.25\textwidth]{imagesMetaTracts/angle-diff.png}
%\caption{Angle between the local orientation of two cylinders that are part of the same MetaTract.}
%\label{fig:angle-diff}
%\end{figure}
The order of the candidate vertices is based on the following characteristics:

@@

In this particular example in Figure~\ref{fig:algo}B we select the orange grid vertex next and repeat the process. We do not select the purple vertex (Figure~\ref{fig:algo}C) because it is not a candidate vertex. If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices and will be of short length. Short tracts are then removed. The MetaTracts generated are shown in Figure~\ref{fig:length_distribution}A. The MetaTracts are colored with the mean orientation direction mapped to the RGB space. That consistent orientation is a key intrinsic feature in the data becomes visually pronounced.
We apply uniform, dense seeding to the XCT volume data to trace and generate the tracts of fiber bundles.

@@

\begin{figure}[tb]
  \centering  	
 %\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/orientation_cluster.PNG}
  \includegraphics[width=0.45\textwidth]{imagesMT2014/image6}
  	\caption{(B) Result of K means clustering with data points projected to the top three eigenvectors as major axes. (A) MetaTracts belonging to orientation cluster 1. (B) MetaTracts belonging to orientation cluster 2. MetaTracts in gray(A,C) show context. }
  \label{fig:orientation_clustering}
  \end{figure}

@@

The directed Hausdorff distance is very sensitive to outliers in the data.
However, because MetaTracts are constructed using cylinders with similar orientations, they are not plagued by outliers.
To cluster based on MetaTract proximity, we used single linkage hierarchical clustering.
Hierarchical clustering has a single parameter $h$, the desired number of clusters.
In hierarchical clustering each objects starts in its own cluster and clusters are merged based on some criterion.
Clusters are merged until there are only $h$ clusters left.
Hierarchical clustering is intuitive since it is easy to trace how clusters are formed and merged.
Single linkage clustering finds pairs of objects $p \in P$ and $q \in Q$ where $P \neq Q$ which are closer than other such pairs, and merges the containing clusters $P$ and $Q$.
We found that single linkage hierarchical clustering had two major drawbacks.
First, the clustering would produce some small clusters of just a few MetaTracts.
These MetaTracts were anomalies caused by noise in the data and did not represent true fiber bundles.
Second, if two fiber bundles were parallel for some of their length and then separated, they would sometimes be clustered into the same fiber bundle.
A short MetaTract which was parallel to both and did not extend into the separation region could form a link between the two fiber bundles, causing them to be clustered into a single bundle.
To address the problem of small clusters, we applied hierarchical clustering and then identified small clusters with few MetaTracts.
We removed the MetaTracts that were in those clusters from the data set and reapplied hierarchical clustering.
To address the problem of short MetaTracts joining different fiber bundles, we applied hierarchical clustering and then removed the
shortest fibers (fibers less than 0.6 times the median length) in each bundle. We then reapplied hierarchical clustering.
We repeated both until a steady state of clusters is reached and no new small fibers can be removed. The results of clustering the orientation clusters are shown in Fig.~\ref{fig:crop-16-decomp}.
\begin{figure*}
\centering
		\includegraphics[width=0.4\textwidth]{imagesMT2014/garb1}~
		\includegraphics[width=0.4\textwidth]{imagesMT2014/image7b}
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
	\caption{(A) Orientation wise decomposition of data set Figure~\ref{fig:data-char}A into the extracted bundles and (B) all of them combined.(C) Shows the results of voxelization of (B). (D) shows a 2D slice of the volume. (E,F) shows the extracted meshes.}
	\label{fig:crop-16-decomp}
\end{figure*}

>>>>

Typically,  carbon fiber reinforced polymer components, more specifically CFRP laminates consists of two main components:

The matrix on the one side, which acts as a bonding component and the reinforcements on the other side, which allows for achieving the desired characteristics. To manufacture CFRP laminates with endless carbon fibers, various production processes are in industrial use. Most of these processes start with the reinforcement component by weaving individual carbon fiber bundles (yarn) into sheets of carbon fiber cloth in a predefined pattern. These sheets of woven carbon fiber cloth, which form the geometrical structure of the final CFRP materials, are also refereed to as fabric. Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth. Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the strength, durability and stiffness properties. Next, resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component. The main function of this matrix component is to act as a bonding component between the individual carbon fiber bundles. For the matrix of CFRP material systems with endless carbon fibers typically resins such as epoxy are frequently used. In special applications additional filler materials, such as colors or other particles, may be integrated. 
%The production process CFRP laminate is finished after curing the resin and the CFRP laminate is ready for further processing.

@@

While fiber bundles are now understood as highly important in determining component properties, the tools for visualization of the internal structure have not developed at the same pace.
To the best of our knowledge, there is no current work that can resolve simple queries such as:
\begin{itemize}
\item{ How to extract and visualize the geometric structure of a particular fiber bundle?}
\item{ How to visualize the interaction between a particular pair of fiber bundles (crossovers/braiding) or a unit cell?}
\item{ Which fiber bundles show a particular orientation? }
\item{ Which fiber bundles are of the same type of yarn (i.e. which bundles show similar sizes or diameters), which is the largest or smallest fiber bundle?}
\end{itemize}

@@

\subsection {Fiber Clustering in Diffusion Tensor Imaging}
\label {subsec:fiberClus}
%!!!!! TODO REFINE FIBER CLUSTERING !!!!!
%The next step is to identify  anatomically relevant fiber bundles. This is achieved through fiber clustering.
The general assumption of fiber clustering in diffusion tensor imaging is that a similarity measure based on factors such as proximity between fibers can be used to interpret a collection of fiber tracts as bundles. Extensive research has been done on automatic DTI fiber clustering methods
% (see \cite{Corouge2004},~\cite{Brun2003},~\cite{Brun2004},~\cite{Zhang2008},~\cite{westinMEDIA02}).

@@

Figure \ref{fig:data-char} shows dataset1 with woven fiber bundles. Dataset1 is 450x300x500 voxels in size with isotropic resolution of $2\mu m$, the datatype is uint8. It clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern. The unit cell is used as a basis structure for CFRP manufacturing. The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material. Figure~\ref{fig:data-char}(A) shows a volume rendering of the dataset. Fig.~\ref{fig:data-char}(B) shows 2D slices along the X and Fig.~\ref{fig:data-char}(C) in Z(left) axis. Figure~\ref{fig:data-char}(D) shows the zoomed in version of the green square. 
The resolution of dataset1 renders the individual carbon fibers in a fiber bundle are hard to resolve using 2D slice images only. Figure~\ref{fig:data-char}(D) contains two bundles going in opposite directions. However, the separation between two fiber bundles is barely visible. The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization. The fiber bundles themselves may differ in terms of the amount of fibers in the bundle, figure~\ref{fig:data-char}(A) shows the large variation in cross section sizes among the bundles.
Depending on the weaving pattern the fiber bundles cross each other in different orientations. In general the number of orientations is limited by the weaving pattern. In all our data sets the number of separate orientations was two. Nevertheless, the weaving pattern may cause the individual bundles to be curved. In consequence of the weaving pattern of the fiber bundles, individual fibers may be adjacent in euclidean distance but belong to different bundles. 
We make the following assumptions on our data:
\begin{itemize}
\item The structure embedded in the data contains fiber bundles of indiscernible fibers.
\item (Local orientation) Each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
\item The local orientation may change gradually inside the fiber bundle.
\item The local orientation may be noisy and not reliable.
\item (Connectivity) Moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
\item Fiber bundles going in different directions only interact near the surface of contact.
\end{itemize}
%\begin{figure}[htp]
%  \centering
%    \subfloat[Overview of the dataset]{\label{figur:3}\includegraphics[scale=0.3]{imagesMetaTracts/crop13-a.eps}}
%  \subfloat[X Axis]{\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/xaxis.eps}}\label{figur:1}\\
%  \subfloat[Y Axis]{\label{figur:2}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/yaxis.eps}}
%  \subfloat[Z Axis]{\label{figur:3}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/zaxis.eps}}
%  \caption{One of our datasets.Fig \ref{fig:orig_data}(a) provides an overview of the entire volume, the rest show slices of each of the major axis.}\label{fig:orig_data}
%\end{figure}

@@

\begin{figure}[tb]
\centering
%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
\includegraphics[width=0.5\textwidth]{imagesMT2014/image3b.png}
\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
\label{fig:flowchart}
\end{figure}

Figure \ref{fig:flowchart} shows our pipeline. The approach consists of multiple steps. MetaTracts generations Sec.~\ref{sec:approach} and  finding plausible fiber bundles (Section~\ref{subsec:fiber-bundles}).
%\subsection {Extracting MetaTracts}
%\label{subsec:ExFiberCyl}
%We explain the different steps of MetaTracts extraction.

@@

\subsection {MetaTracts generation}
\label{subsec:MetaTracts-generation}
In $\mathbb{R}^2$ all the above properties hold except that cylinders are replaced by rectangles. We explain the process of MetaTracts generation in $\mathbb{R}^2$. The procedure trivially extends to $\mathbb{R}^3$. The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a $R_H$[0,1]. In Figure \ref{fig:algo} all regions which have reliable Hessian less than the threshold have been marked in blue. Let the seed point associated with the MetaTract be grid point $C_{p}$ (Property 2, Section~\ref{subsec:mtprop}). The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the dark green arrow (Figure~\ref{fig:algo}C). We generate a rectangle of length L and radius R, the green dashed rectangle shows the rectangle generated at $C_p$ (Property 3 and 5, Section~\ref{subsec:mtprop}). 
\begin{figure}[tb]
  \begin{center}
   \includegraphics[ width=0.45\textwidth]{imagesMT2014/image4}
 % \includegraphics[ width=0.45\textwidth]{imagesMT2014/MT_abstract} 
  \end{center}
  \vspace{-1em}
 \caption{MetaTracts generation: (B) MetaTract generation, region with $R_H$ less than threshold. (A) close up of marked region in (B). (C) Parts of a single MetaTract.}
  \label{fig:algo}
    \vspace{-1.5em}
\end{figure}
The set of vertices which are in the green region but not overlapped by blue are possible candidates for the start point of the next cylinder in a particular MetaTract. We call these vertices \textit{``candidate vertices"}. From these we select another grid vertex which will be the start point for the next cylinder.
%\begin{figure}
%\centering
% \includegraphics[width=0.25\textwidth]{imagesMetaTracts/angle-diff.png}
%\caption{Angle between the local orientation of two cylinders that are part of the same MetaTract.}
%\label{fig:angle-diff}
%\end{figure}
The order of the candidate vertices is based on the following characteristics:

@@

In this particular example in Figure~\ref{fig:algo}(B) we select the orange grid vertex next and repeat the process. We do not select the purple vertex (Figure~\ref{fig:algo}(C)) because it is not a candidate vertex. If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices and will be of short length. Short tracts are then removed. The MetaTracts generated are shown in Figure~\ref{fig:length_distribution}A. The MetaTracts are colored with the mean orientation direction mapped to the RGB space. That consistent orientation is a key intrinsic feature in the data becomes visually pronounced.
We apply uniform, dense seeding to the XCT volume data to trace and generate the tracts of fiber bundles.

@@

\begin{figure}[tb]
  \centering  	
 %\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/orientation_cluster.PNG}
  \includegraphics[width=0.55\textwidth]{imagesMT2014/garb2}
  	\caption{(B) Result of K means clustering with data points projected to the top three eigenvectors as major axes. (A) MetaTracts belonging to orientation cluster 1. (B) MetaTracts belonging to orientation cluster 2. MetaTracts in gray(A,C) show context. }
  \label{fig:orientation_clustering}
  \end{figure}

@@

The directed Hausdorff distance is very sensitive to outliers in the data.
However, because MetaTracts after orientation clustering are constructed using cylinders with similar orientations, they are not plagued by outliers.
To cluster based on MetaTract proximity, we used single linkage hierarchical clustering.
Hierarchical clustering has a single parameter $h$, the desired number of clusters.
In hierarchical clustering each objects starts in its own cluster and clusters are merged based on some criterion.
Clusters are merged until there are only $h$ clusters left.
Hierarchical clustering is intuitive since it is easy to trace how clusters are formed and merged.
Single linkage clustering finds pairs of objects $p \in P$ and $q \in Q$ where $P \neq Q$ which are closer than other such pairs, and merges the containing clusters $P$ and $Q$.
We found that single linkage hierarchical clustering had two major drawbacks.
First, the clustering would produce some small clusters of just a few MetaTracts.
These MetaTracts were anomalies caused by noise in the data and did not represent true fiber bundles.
Second, if two fiber bundles were parallel for some of their length and then separated, they would sometimes be clustered into the same fiber bundle.
A short MetaTract which was parallel to both and did not extend into the separation region could form a link between the two fiber bundles, causing them to be clustered into a single bundle.
To address the problem of small clusters, we applied hierarchical clustering and then identified small clusters with few MetaTracts.
We removed the MetaTracts that were in those clusters from the data set and reapplied hierarchical clustering.
To address the problem of short MetaTracts joining different fiber bundles, we applied hierarchical clustering and then removed the
shortest fibers (fibers less than 0.6 times the median length) in each bundle. We then reapplied hierarchical clustering.
We repeated both until a steady state of clusters is reached and no new small fibers can be removed. The results of clustering the orientation clusters Fig.~\ref{fig:orientation_clustering}(A,C) are shown in Fig.~\ref{fig:orientation_clustering}(D,E) respectively. Note how the purple cluster Fig.~\ref{fig:orientation_clustering}(E) bundle ``7'' is extracted well.
\begin{figure}
\centering
		%\includegraphics[width=0.4\textwidth]{imagesMT2014/garb1}~
		\includegraphics[width=0.5\textwidth]{imagesMT2014/vol}
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
	\caption{(A) Orientation wise decomposition of data set Figure~\ref{fig:data-char}A into the extracted bundles and (B) all of them combined.(C) Shows the results of voxelization of (B). (D) shows a 2D slice of the volume. (E,F) shows the extracted meshes.}
	\label{fig:crop-16-decomp}
\end{figure}

### 2014-09-15 metaTracts-pvis2014.tex
2nd pass
updated text

For the purpose of characterizing larger areas of CFRP materials such as complete unit cells integrating the recurring weaving pattern, the proposed workflow is designed to handle XCT scans of low resolution, in which the individual fibers of the bundles to be extracted are not or just barely visible.  
The proposed workflow applies and extends methods originally presented for the field of diffusion tensor imaging in order to cope with the low data quality: 
First, a coarse version of integral curves is used to trace subsections of the individual fiber bundles in the woven CFRP materials. We call these sections MetaTracts.
In the second step these extracted fiber bundle sections (MetaTracts) are clustered using a two step approach, wherein the first step clusters according to orientation and the second step cluster by proximity of the MetaTracts. 
Finally, the tool generates volumetric representations as well as surface models of the extracted fiber bundles to be exported for further analysis.

>>>>

For the purpose of characterizing larger areas of CFRP materials such as complete unit cells integrating the recurring weaving pattern, the proposed workflow is designed to handle XCT scans of low resolution, in which the individual fibers of the bundles to be extracted are not or just barely visible.  
The proposed workflow applies and extends methods originally presented for the field of diffusion tensor imaging in order to cope with the low data quality: 
First, a coarse version of integral curves is used to trace subsections of the individual fiber bundles in the woven CFRP materials. We call these sections MetaTracts.
In the second step these extracted fiber bundle sections (MetaTracts) are clustered using a two step approach, wherein the first step clusters according to orientation and the second step clusters by proximity of the MetaTracts. 
Finally, the tool can also  generate volumetric representations as well as surface models of the extracted fiber bundles to be exported for further analysis.
